Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\

### 2018-06-04 Presentations/Paper/paper_thread_pool.tex
Continue writing implementation section of paper

\begin{figure*}[htbp]
	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
	\caption{Design of the thread pool system}
	\label{fig0}
\end{figure*}

@@

\subsection{Thread Pool and Task Queue}
Threads and tasks are two major entities in the thread pool system. As we can observe from Figure 1, thread pool and task queue are two data structures used to store the required information related to the threads and the submitted tasks, respectively. Each spawned thread can be either working on a task or waiting for the next task. Consequently, we define the states of a thread to \emph{busy} and \emph{waiting}.

\subsection{Enqueueing Tasks}
\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_queue.png}
	\caption{Workflow of task enqueueing and execution in the thread pool}
	\label{fig1}
\end{figure}

\subsection{Waiting for Tasks}
\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/waitingconcept.png}
	\caption{Concept of waiting for tasks using a slotmap}
	\label{fig2}
\end{figure}
As mentioned in the previous section, ids are assigned to the tasks before they are enqueued.

>>>>

\begin{figure*}[htbp]
	\includegraphics[width=1.0\textwidth]{img/pool_structure.png}
	\caption{Design of the thread pool system TODO: Fix Image!}
	\label{fig0}
\end{figure*}

@@

\subsection{Thread Pool and Task Queue}
Threads and tasks are two major entities in the thread pool system. As we can observe from Figure 1, thread pool and task queue are two data structures used to store the required information related to the threads and the submitted tasks, respectively. Each spawned thread can be either working on a task or waiting for the next task. Consequently, we define the states of a thread as \emph{busy} and \emph{waiting}. A busy thread changes to waiting after processing its task if no other task is enqueued. Waiting tasks do busy waiting until there are new tasks available.

Each assigned task contains a function pointer to a routine and a pointer to data. We implemented the task queue as a generic priority queue that uses a mutex thread to enqueue tasks threadsafe. Every time a new task is enqueued to the task queue, two scenarios can occur. If the new task has a priority, it is placed in the right slot of the task queue regarding its priority and the subsequent tasks are relocated. If the  new task has no priority, it is appended to the end of the queue. Furthermore, tasks can refer to a handle, which is used by the thread pool to wait for a specific amount of tasks.

\subsection{Processing Tasks}
\textbf{TODO: maybe table of provided functions} \\
We define the processing of tasks as a procedure of 3-4 steps. In Figure 2, we show the flow of task enqueueing and execution in the thread pool system. In the first step, tasks are passed as a parameter to the enqueue function provided by the thread pool. The tasks have to be created before passing and later on freed after processing by the calling function.  We provide different functions to enqueue a single task or an array of tasks.

In the next step, the tasks will be enqueued into the priority queue of the thread pool. Optionally, a handle can be passed with the tasks or generated by the thread pool to force the calling function to wait for the tasks. After enqueueing the tasks to the priority queue, the waiting threads in the thread pool pop the enqueued tasks from the priority queue and executes the routine of the task. This happens asynchronously to the calling function and therefore is not considered as an own step of the process.

The third step is only executed if the tasks refer to a valid handle and waiting is enforced by the calling function. This can be applied either by executing a provided waiting function from the thread pool or by calling \emph{thread\_pool\_enqueue\_tasks\_wait}. We will present our design of waiting for a specific amount of tasks in Section 3.D. The last step only contains the return to the calling function.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_queue.png}
	\caption{Workflow of task enqueueing and execution in the thread pool}
	\label{fig1}
\end{figure}

\subsection{Waiting for Tasks}
\begin{figure}
	\includegraphics[width=0.4\textwidth]{img/waitingconcept.png}
	\caption{Concept of waiting for tasks using a slotmap}
	\label{fig2}
\end{figure}
As mentioned in the previous section, ids are assigned to the tasks before they are enqueued.
